---
title: "HW1 Basics"
author: "Wang, Yixing"
date: "`r format(Sys.time(), '%d %h, %Y, %I:%M %p')`"
output: html_document
---


1. Insert Your codes
2. Show your answers outside the chunk.
3. Knit your HW and submit the html version.

# 1 Introductory

1. What is the name, number and date of the installed R  version on your computer?

ans: R version 4.5.1 (2025-06-13)

2. What is the number and date of the installed R studio version on your computer?

ans:2025.05.1+513

3. The function `sd` calculates the standard deviation. Calculate the standard deviation
of the numbers 0 ,1 ,2, ..., 1000. 


```{r}
#insert your code, for example
z <- 0:1000
sd(z)
```

ans: 289.1081

4. Using R, calculate $$ \frac{7x+23}{5^{6.2-5.1}} $$ when $x=-3.1$


```{r}
#insert your code
x <- -3.1
ans <- (7 * x + 23) / (5^(6.2 - 5.1))
print(ans)
```
ans:0.2213484

5. Using R, calculate the square root of the average of the numbers 234.5, 6.07, 88.44, and `sin (1/8) `


```{r}
#insert your code
x<-c(234.5,6.07,88.44,sin(1/8))
x
mean(x)
sqrt(mean(x))
```

ans:9.071035

6. Find `ln 2.5`


```{r}
#insert your code
ans <- log(2.5)
print(ans)
```

ans:0.9162907

7. Create an object `a1` that stores $$3^{-5} \times 4^{4}\times 6^{1/6} $$



```{r}
#insert your code
a1 <- 3^(-5) * 4^4 * 6^(1/6)
print(a1)
```

ans:1.420122


8. Create a new object `a2` with the value $$ -9.3 \times 5^{-12}$$


```{r}
#insert your code
a2 <- -9.3 * 5^(-12)
print(a2)
```

ans:-3.80928e-08

9. Print the values of `a1, a2` and their product without using the print command. 



```{r}
#insert your code
a1 <- 1.420122
a2 <- -3.80928e-08
a1
a2
a1 * a2
```



10. Try the commands `pi, round(pi), round(pi, digits=4),  trunc(pi), exp(pi*1i)`. print out 16 digits of pi(use `options()` function)



```{r}
#insert your code
pi
round(pi)
round(pi, digits=4)
trunc(pi)
exp(pi*1i)
options(digits=16)
pi
```

ans:

11. Which of the five numbers 0, Inf, -Inf, NaN, and NA are considered not missing?

ans: 

12. Assign the numbers 1, 2, 3, ..., 1000 to a variable x. Calculate the inverse tangent of
the reciprocal of x,  and assign it to a variable y. print out the sum of y. (Do not print y or x)



```{r}
#insert your code
x <- 1:1000
y <- atan(1 / x)
print(sum(y))

```

ans:7.209896

13. Which number is larger: $e^\pi$ or $\pi^e$ ?. Use R to compare them.


```{r}
#insert your code
x <- exp(pi)
y <- pi^exp(1)
x
y
x>y

```

ans:23.14069
 22.45916
  TRUE

14. Your friend Pedro wants to print the string `Math 3700` to the console. He enters the following command but encounters an error. What went wrong? What should he have done instead?

`print(Math 3700)`

`Error: unexpected numeric constant in "print(Math 3700"`

ans:Text data  must be enclosed in quotation marks.
He should have enclosed the entire string in quotes like:
print("Math 3700") 

15. What is the result of `NA + 5`? Why do you think that is?

ans:The result is NA.
Any arithmetic operation involving (NA) results in a missing value. 