---
title: 'HW2: Vectors'
author: "Wang, Yixing"
date: "2025.9.18 "
output: html_document
---


# 2 Vectors

1. How would you create a vector `V` containing the values 0, 0.2,0.4,0.6,0.8, and 1, using the `seq` function?

```{r}
#insert your code
V <- seq(from = 0, to = 1, by = 0.2)
print(V)
```

2. Change the value 0.4 to -0.4 in vector V. Use bracket assignment

```{r}
#insert your code
V <- seq(from = 0, to = 1, by = 0.2)
V[3] <- -0.4
print(V)
```


3. Name the elements of `V`: January, February, ...

```{r}
#insert your code
V <- month.name
print(V)
```


4. The weights of five people before and after a diet programme are given in the table.

|Before |78| 72| 78| 79| 105|
|-------|--|--|--|--|--|--|
|After  |67| 65| 79| 70| 93|



Read the `before` and `after` values into two different vectors called before
and after. Use R to evaluate the amount of weight lost for each participant.
What is the average amount of weight lost?


```{r}
#insert your code
before <- c(78,72,78,79,105)
after <- c(67,65,79,70,93)
weight_lost <- before - after
print(weight_lost)
average_lost <- mean(weight_lost)
print(average_lost)
```


5. The function `rnorm()` generates normal random variables. For instance, `rnorm(10)` gives a vector of 10 i.i.d. standard normals. Generate 20 standard normals, and store them as x. Then obtain subvectors
of

(i) the entries in x which are less than 1;

(ii) the entries between -1/2 and 1;
(iii) the entries whose absolute value is larger than 1.5.


```{r}
# insert your code
x <- rnorm(20)
x
subvector_i <- x[x < 1]
subvector_i
subvector_ii <- x[x > -0.5 & x < 1]
subvector_ii
subvector_iii <- x[abs(x) > 1.5]
subvector_iii
```



6. Repeat the vector `c(1,2,5, 7, 13)` twice, with each element repeated 7 times, and store the result in `xrep`. 
Display the result sorted from largest to smallest.


```{r}
#insert your code
xrep <- rep(c(1, 2, 5, 7, 13), each = 7, times = 2)
print(xrep)
```


7. Use the data set `airquality`. Find 


  a. the highest and lowest temperature

  b. Number of days with temperature more than 80 


```{r}
#insert your code
air <- airquality
highest_temp <- max(air$Temp, na.rm = TRUE)
lowest_temp <- min(air$Temp, na.rm = TRUE)
print(highest_temp)
print(lowest_temp)
hot_days_count <- sum(air$Temp > 80, na.rm = TRUE)
print(hot_days_count)
```

8. Use the function `paste`  and `:` operator to create the following character vector called `ID25`:
 `("ID 1", "ID 2", ....., "ID 25")`

```{r}
#insert your code
ID25 <- paste("ID", 1:25)
print(ID25)
``` 